---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.path = "man/figures/README-",
  echo = TRUE,
  fig.width = 8,
  fig.height = 6
)
```

# trendbreaker

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3888494.svg)](https://doi.org/10.5281/zenodo.3888494)
[![R build status](https://github.com/reconhub/trendbreaker/workflows/R-CMD-check/badge.svg)](https://github.com/reconhub/trendbreaker/actions)
<!-- badges: end -->

The *trendbreaker* package implements tools for detecting changes in temporal
trends of a single response variable. It implements the **A**utomatic 
**S**election of **M**odels and **O**utlier **De**tection for **E**pidemmics 
(ASMODEE), an algorithm originally designed for detecting changes in COVID-19 
case incidence. 

ASMODEE proceeds by:

1. defining a training set excluding the last *k* data points
2. identifying the temporal trend in the training set by fitting a range of
   (user-specified) models to the data and retaining the best predicting /
   fitting model
3. calculating the prediction interval (PI) of the temporal trend
4. classifying any data point outside the PI as outlier

The algorithm can be applied with fixed, user-specified value of *k*, so as to
monitor potential changes in this recent time period. Alternatively, the optimal
value of *k* can be determined automatically.

**Disclaimer:** this is work in progress. Please reach out to the authors before
using this package. Also note this package may soon be renamed to avoid clashes
with other projects and reflect a more general scope.


## Getting started

Once it is released on [CRAN](https://CRAN.R-project.org), you will be able to
install the stable version of the package with:

``` r
install.packages("trendbreaker")
```

The development version can be installed from [GitHub](https://github.com/) with:

``` r
if (!require(remotes)) {
  install.packages("remotes")
}
remotes::install_github("reconhub/trendbreaker")
```

The best place to start for using this package is to read the documentation of
the function `asmodee` and run its example:

```{r eval = FALSE}
library(trendbreaker)
?asmodee
example(asmodee)
```


## Main features

The package implements the following main functions

* `asmodee`: implements the Automatic Selection of Models and Outlier DEtection
  for Epidemics
  
* `detect_changepoint`: a function to detect the points at which recent data
  deviate from previous temporal trends using a fitted model and
  data

* `detect_outliers`: a function to identify outliers using a fitted model and
  data
  


### ASMODEE

We illustrate ASMODEE using publicly available NHS pathways data recording
self-reporting of potential COVID-19 cases in England 
(see `?nhs_pathways_covid19` for more information). 

```{r asmodee}
library(trending)     # for model interface
library(trendbreaker) # for ASMODEE
library(dplyr)        # for data manipulation

# load data
data(nhs_pathways_covid19)

# select last 6 weeks of data
first_date <- max(nhs_pathways_covid19$date, na.rm = TRUE) - 6*7
pathways_recent <- nhs_pathways_covid19 %>%
  filter(date >= first_date)

# define candidate models
models <- list(
  regression = lm_model(count ~ day),
  poisson_constant = glm_model(count ~ 1, family = "poisson"),
  negbin_time = glm_nb_model(count ~ day),
  negbin_time_weekday = glm_nb_model(count ~ day + weekday)
)

# analyses on all data
counts_overall <- pathways_recent %>%
  group_by(date, day, weekday) %>%
  summarise(count = sum(count))

# results with automated detection of 'k'
res <- asmodee(counts_overall, models, method = evaluate_aic)
res
plot(res, "date")
```

ASMODEE would typically be more useful to investigate shifts in temporal trends
from a large number of time series (e.g. at a fine geographic scale).  To make
this sort of analysis easier *trendbreaker* also works with 
[*incidence2*](https://github.com/reconhub/incidence2/) objects.  To illustrate
this we can consider trends over NHS regions.

```{r incidence2}
library(incidence2)   # for `incidence()` objects

# select last 6 weeks of data
first_date <- max(nhs_pathways_covid19$date, na.rm = TRUE) - 6*7
pathways_recent <- filter(nhs_pathways_covid19, date >= first_date)

# create incidence object with extra variables
lookup <- select(pathways_recent, date, day, weekday) %>%  distinct()

dat <-
  pathways_recent %>%
  as_incidence(date_index = date, counts_var = count, group_vars = nhs_region) %>%
  left_join(lookup, by = "date")

# define candidate models
models <- list(
  regression = lm_model(count ~ day),
  poisson_constant = glm_model(count ~ 1, family = "poisson"),
  negbin_time = glm_nb_model(count ~ day),
  negbin_time_weekday = glm_nb_model(count ~ day + weekday)
)

# analyses on all data
res <- asmodee(dat, models, method = evaluate_aic)

plot(res, "date")
```
